Exact asymptotic expressions for local clustering in hyperbolic random graphs

### 2018-11-16 concentration_argument.tex
typos in the proof of thm 1.1. in section 2 and typos in section 3

\section{Concentration of heights for nodes with degree $k$}\label{sec:concentration_argument}

Many of the computations in this paper will involve integrals whose integrand contains $\Prob{\Po(\mu_n(y)) = k}$, where $\Po(\lambda)$ denotes a Poisson random variable with mean $\lambda$ and $\mu_n(y) \approx e^{y/2}$. For instance, the expected adjusted local clustering function $c_\infty(k)$ in $G_\Pcal$ equals
\[
	\frac{\int_0^\infty \Prob{\Po(\Mu{\BallPo{0,y}}) = k} \Delta_{\Pcal}(y) \alpha e^{-\alpha} \dd y}
	{\int_0^\infty \Prob{\Po(\Mu{\BallPo{0,y}}) = k} \alpha e^{-\alpha} \dd y},
\]
where $\Mu{\BallPo{0,y}} = \xi_{\alpha,\nu} e^{y/2}$.

@@

Recall that $\rho(y, k)$ is the probability density function of a Poisson random variable with mean $\mu_{\alpha, \nu}(B_\Pcal(y)) = \xi_{\alpha,\nu}e^{\frac{y}{2}}$. 
For any sequence $\{k_n\}_{n \ge 1}$, possibly constant, define
\begin{equation}\label{eq:def_kappa_n}
	\kappa_n := \begin{cases}
		\log(n) &\mbox{if } k_n = \bigT{1},\\
		\sqrt{k_n \log(k_n)} &\mbox{else.}
	\end{cases}
\end{equation}
and define further, for any $C > 0$
\begin{equation}\label{eq:def_K_C_set}
	\Kcal_C(k_n) = \left\{p \in \R : \frac{k_n - C \kappa_n}{\xi_{\alpha,\nu}} \vee 0 \le e^{\frac{y}{2}}
	\le \frac{k_n + C \kappa_n}{\xi_{\alpha,\nu}} \wedge e^{R_n/2} \right\},
\end{equation}
Note that if $k_n = \Omega(\log(n))$ then $e^{y/2} = \bigT{k_n}$ whenever $p \in \Kcal_C(k_n)$. The next lemma states that for a large class of functions $h(y)$, to compute the integral 
\[
	\int_{\Rcal_n} \rho(y,k_n) h(y) f_{\alpha,\nu}(x,y) \dd x \dd y
\]
it is enough to consider integration of $\Kcal_C(k_n)$ instead of $\Rcal_n$. More precisely, for any $\ell_n = (1 + \smallO{1})k_n$ it is enough to consider $\Kcal_C(\ell_n)$. The flexibility of using $(1+\smallO{1})k_n$ instead of $k_n$ will prove useful in later sections.

\begin{lemma}\label{lem:concentration_argument}
Let $\alpha > \frac{1}{2}$, $\nu > 0$, $\{k_n\}_{n \ge 1}$ be any positive sequence such that $k_n = o(n^{\frac{1}{2\alpha + 1}})$ and let $\ell_n = k_n(1 + \epsilon_n)$, with $\epsilon_n \to 0$. In addition let $\beta < \alpha$ and $h : \R_+ \mapsto \R$ be a any function such that $h(y) = g(y)e^{\beta y}$ with $|g(y)|$ uniformly bounded on $\R_+$. Then, if we define for $C > 0$
\[
	\lambda_n^\pm = \ell_n \pm C \kappa_n, 
	\quad a_n^\pm = 2 \log\left(\frac{\lambda_n^\pm}{\xi_{\alpha,\nu}}\right),
\] 
we have
\begin{equation}\label{eq:error_bound_int_rho_not_K}
\begin{aligned}
	\int_{\R_+ \setminus [a_n^-, a_n^+]} \rho(y,k_n) h(y) \alpha e^{-\alpha y} \dd x \dd y
	&= \begin{cases}
		\bigO{\log(n)^{k_n} n^{-\frac{C}{2}}} &\mbox{if } k_n = \bigT{1} \\
		\bigO{k_n^{-(1+C^2)/2}} &\mbox{else.}
	\end{cases}
\end{aligned}
\end{equation}
as $n \to \infty$.

@@

\subsection{Average degree in the Hyperbolic graph}\label{ssec:average_degree_HP_n}

Recall that under the coupling between the hyperbolic random graph and the finite box model $|x-x^\prime|_{\pi e^{r_n/2}} \le \Omega(r,r^\prime)$, while the coupling lemma (Lemma~\ref{lem:asymptotics_Omega_hyperbolic}) gives that  
\[
	e^{\frac{1}{2}(y+y^\prime)} - K e^{\frac{3}{2}(y+y^\prime) - R_n} \leq \Omega(r, r^\prime) 
		\leq  e^{\frac{1}{2}(y+y^\prime)} + K e^{\frac{3}{2}(y+y^\prime) - R_n},
\]
for $y + y^\prime < R_n$. This result enables us to determine the measure of a ball around a given point $p=(0,y)$ which will be fairly useful in our subsequent analysis.

@@

\begin{proof}
We will split the computation of $\Mu{\BallHyp{0,y}}$ over the case $y^\prime > R_n - y$ and $y^\prime \le R_n - y$ where for the latter we utilize Lemma~\ref{lem:asymptotics_Omega_hyperbolic}. Recall that $\Mu{\BallPo{0,y}} = \xi_{\alpha,\nu} e^{y/2}$ where $\xi_{\alpha,\nu} = 4\nu \alpha/(2\alpha - 1)\pi$.

@@

Next we will establish upper and lower bounds on $\Mu{\BallHyp{(0,y)} \cap\Rcal_n[(0,R_n-y)]}$. Using Lemma~\ref{lem:asymptotics_Omega_hyperbolic} we have 
\begin{align*}
	\Mu{\BallHyp{(0,y)} \cap\Rcal_n[(0,R_n-y)]} 
	&\le \frac{2\nu \alpha}{\pi} \int_{0}^{R_n - y} \left(e^{\frac{y + y^\prime}{2}} + K e^{\frac{3}{2}(y + y^\prime) - R_n}\right)
		e^{-\alpha y^\prime} \dd y^\prime\\
	&= \Mu{\BallPo{0,y}}\left(1 - e^{-(\alpha - \frac{1}{2})(R_n - y)}\right) \\
	&\hspace{10pt}+ \frac{2\nu \alpha}{\pi} K e^{\frac{3 y}{2} - R_n}\int_0^{R_n-y} e^{(\frac{3}{2} - \alpha)y^\prime} \dd y^\prime
\end{align*}
The last integral depend on the value of $\alpha$,
\[
	\int_0^{R_n-y} e^{(\frac{3}{2} - \alpha)y^\prime} \dd y^\prime
	= \begin{cases}
		\frac{2}{3 - 2\alpha}\left(e^{(\frac{3}{2} - \alpha)(R_n - y)} - 1\right) &\mbox{if } 1/2 < \alpha < 3/2,\\
		R_n - y &\mbox{if } \alpha = 3/2,\\
		\frac{2}{2\alpha-3}\left(1 - e^{-(\alpha - \frac{3}{2})(R_n - y)}\right) &\mbox{if } \alpha > 3/2.
	\end{cases}
\]
Therefore we get
\begin{align*}
	&\hspace{-30pt}\frac{2\nu \alpha}{\pi} K e^{\frac{3 y}{2} - R_n}
		\int_0^{R_n-y} e^{(\frac{3}{2} - \alpha)y^\prime} \dd y^\prime\\
	&= \Mu{\BallPo{0,y}}\begin{cases}
		\frac{(2\alpha - 1)K}{3 - 2\alpha}\left(e^{-(\alpha - \frac{1}{2})(R_n - y)} - e^{-(R_n - y)}\right)
		&\mbox{if } 1/2 < \alpha < 3/2,\\
		\frac{(2\alpha -1)K}{2} (R_n - y)e^{-(R_n - y)} &\mbox{if } \alpha = 3/2,\\
		\frac{(2\alpha - 1)K}{2\alpha - 3} \left(e^{-(R_n - y)} - e^{-(\alpha - \frac{1}{2})(R_n - y)}\right)
		&\mbox{if } \alpha > 3/2.
	\end{cases}\\
	&= \Mu{\BallPo{0,y}} \phi_{\H,n}^{(2)}(y)
\end{align*}
and hence
\[
	\Mu{\BallHyp{(0,y)} \cap\Rcal_n[(0,R_n-y)]} = \Mu{\BallPo{0,y}}\left(1 - e^{-(\alpha - \frac{1}{2})(R_n - y)} 
	+ \phi_{\H,n}^{(2)}(y)\right).
\]
Combining this with~\eqref{eq:mu_hyperbolic_ball_part_1} yields the required upper bound.

@@

\begin{lemma}\label{lem:average_degree_P_n}
For all $p \in \Rcal_n$, such that $y > 2\log(\pi/2)$,
\[
	\mu_{\alpha,\nu}(B_{\Pcal,n}(p)) = \mu_{\alpha,\nu}(B_\Pcal(p))\left(1 - \phi_n(y)\right)
\]
where $\phi_n(y) \ge 0$ is given by
\[
	\phi_n(y) = \left(\frac{\pi}{2}\right)^{-(2\alpha - 1)}e^{-(\alpha-\frac{1}{2})(R_n - y)}
	- \frac{(2\alpha - 1)\pi}{4\alpha}\left(\left(\frac{\pi}{2}\right)^{-2\alpha} 
	e^{-(\alpha - \frac{1}{2})(R_n - y)} - e^{-(\alpha - \frac{1}{2})R_n - \frac{y}{2}}\right).
\]
If, on the other hand, $y \le 2 \log(\pi/2)$ then
\[
	\mu_{\alpha,\nu}(B_{\Pcal,n}(p)) = \mu_{\alpha,\nu}(B_\Pcal(p))\left(1 - e^{-(\alpha - \frac{1}{2})R_n}\right).
\]
\end{lemma}

@@

\subsection{Concentration argument for Hyperbolic and finite box models}

With the results on the average degrees in the two finite Poisson model, we can now establish a result that allows us to extend the concentration argument to the case where instead of $\rho(y,k)$ we consider the degree distributions $\rho_{\HP,n}(y,k)$ and $\rho_{n}(y,k)$ in $G_{\HP,n}$ and $G_{\Pcal,n}$, respectively. The lemma is stated in a more general form, to make it applicable to slightly other cases later in the paper. To understand the general conditions presented in the statement of the lemma we write $\mu_n(y) = \Mu{\BallPon{y}}$ and $\mu(y) = \Mu{\BallPo{y}}$ and recall the result from Lemma~\ref{lem:average_degree_P_n}
\[
	\mu_n(y) = \mu(y)(1 - \phi_n(y)).
\]
For the error function $\phi_n(y)$ we note that for any $0 < \varepsilon < 1$, as $n \to \infty$, 
\[
	\sup_{0 \le y \le (1-\varepsilon)R_n} \phi_n(y) = \smallO{1},
\]
while
\[
	\frac{\partial \phi_n(y)}{\partial y} = \bigT{\phi_n(y)}.
\]
These are exactly the crucial properties of the error function that allow for a concentration argument. Moreover, if the integrand also contains a function $g_n(y)$ that is reasonably nice, we can replace the $\rho_{n}(y,k_n)$ in the integrand by $\rho(y,k_n)$. In particular, the triangle counting function $\Delta_{\Pcal}(y)$ will satisfy these conditions, see Section~\ref{sec:clustering_Pn_to_P}.

\begin{lemma}\label{lem:concentration_argument_rho_approximation}
Let $\alpha > \frac{1}{2}, \nu > 0$, $k_n$ be any positive sequence such that $k_n = o(n^{\frac{1}{2\alpha + 1}})$, $\ell_n = (1 + \epsilon_n)k_n$, with $\epsilon_n \to 0$ and let $\Kcal_{C}(\ell_n)$ be defined as in~\eqref{eq:def_K_C_set}. In addition, define $\hat{\rho}_n(y,k) = \Prob{\Po(\hat{\mu}_n(y)) = k}$, where $\hat{\mu}_n(y)$ is any function satisfying 
\[
	\Mu{\BallPo{0,y}}(1 - \phi_n^{(1)}(y)) \le \hat{\mu}_n(y) \le \Mu{\BallPo{0,y}}(1 + \phi_n^{(2)}(y))
\]
where $\phi_n^{(i)} : \R_+ \to \R_+$ are such that $\sup_{0 \le y \le (1-\varepsilon)R_n} \phi_n(y)^{(i)} = \smallO{1}$ for any $0 < \varepsilon < 1$ and $\partial \phi_n^{(i)}(y)/\partial y = \bigT{\phi_n^{(i)}(y)}$. Then the following holds from some $C > 0$ large enough:
\begin{enumerate}
\item for any sequence of functions $h_n(y) : \R_+ \mapsto \R$ such that $h_n(y) = \hat{\rho}_n(y,k_n) \bigO{k_n^s e^{\beta y}}$ for some $s < 2\alpha(2\alpha + 1)$ and $\beta < \alpha$, we have,
\[
	\int_{\Rcal_n} h_n(y) f_{\alpha,\nu}(x,y) \dd x \dd y 
	= (1 + o(1)) \int_{\Kcal_{C}(\ell_n)} h_n(y) f_{\alpha,\nu}(x,y) \dd x \dd y.
\]
\item If, in addition, $h_n(y) = \hat{\rho}_n(y,k_n) g_n(y)$ with $g_n$ satisfying $g_n(y(1+\smallO{1}) = (1+\smallO{1})g_n(y)$, we have,
\[
	\int_{\Rcal_n} h_n(y) f_{\alpha,\nu}(x,y) \dd x \dd y 
	= (1 + o(1)) \int_{\Rcal_n} \rho(y,k_n) g_n(y) f_{\alpha,\nu}(x,y) \dd x \dd y.
\]
\end{enumerate}
\end{lemma}

\begin{proof}
Similar to the proof of Lemma~\ref{lem:concentration_argument} we define 
\[
	\lambda_n^\pm = \ell_n \pm C \kappa_n, \quad \text{and} \quad a_n^\pm = 2 \log\left(\frac{\lambda_n^\pm}{\xi_{\alpha,\nu}}\right).
\] 
In addition we fix $\varepsilon > 0$ such that $\varepsilon < \min\{1- \frac{s}{2\alpha(2\alpha + 1)}, 1\}$. Then since $h_n(y) = \hat{\rho}_n(y,k_n) \bigO{k_n^s}$ we have
\begin{align*}
	k_n^s \int_{(1-\varepsilon)R_n}^{R_n} \hat{\rho}_n(y,k_n) e^{-\alpha y} \dd y
	&= \bigO{1} \hat{\rho}_n((1-\varepsilon)R_n,k_n) k_n^s e^{-\alpha(1-\varepsilon)R_n} \\
	&= \bigO{\hat{\rho}_n((1-\varepsilon)R_n,k_n) k_n^s n^{-2\alpha(1 - \varepsilon)}} \\
	&= \smallO{ \hat{\rho}_n((1-\varepsilon)R_n,k_n) n^{\frac{s}{2\alpha + 1} - 2\alpha(1-\varepsilon)}} = \smallO{1},
\end{align*}
where the last step follows by our choice of $\varepsilon$. Hence it is enough to prove both statement on $\Rcal_n((1-\varepsilon)R_n, R_n)$ instead of $\Rcal_n$.
\PvdH{This is where we use that $s < 2\alpha(2\alpha + 1)$. I think this can be removed by using a more careful bound but I am not sure this is needed.}

@@

In particular we conclude that, similar to the infinite limit model, concentrations arguments as described in Remark~\ref{rmk:concentration_argument} can be applied in the case of the hyperbolic random graph and finite box model.

>>>>

\section{Concentration of heights for vertices with degree $k$}\label{sec:concentration_argument}

Many of the computations in this paper will involve integrals whose integrand contains $\Prob{\Po(\mu_n(y)) = k}$, where $\Po(\lambda)$ denotes a Poisson random variable with expectation $\lambda$ and $\mu_n(y) \approx e^{y/2}$. For instance, the expected modified local clustering function $c_\infty(k)$ in $G_\Pcal$ equals
\[
	\frac{\int_0^\infty \Prob{\Po(\Mu{\BallPo{0,y}}) = k} \Delta_{\Pcal}(y) \alpha e^{-\alpha y} \dd y}
	{\int_0^\infty \Prob{\Po(\Mu{\BallPo{0,y}}) = k} \alpha e^{-\alpha y} \dd y},
\]
where $\Mu{\BallPo{0,y}} = \xi_{\alpha,\nu} e^{y/2}$.

@@

Recall that $\rho(y, k)$ is the probability density function of a Poisson random variable with expectation $\mu_{\alpha, \nu}(B_\Pcal(y)) = \xi_{\alpha,\nu}e^{\frac{y}{2}}$. 
For any sequence $\{k_n\}_{n \ge 1}$, possibly constant, define
\begin{equation}\label{eq:def_kappa_n}
	\kappa_n := \begin{cases}
		\log(n) &\mbox{if } k_n = \bigT{1},\\
		\sqrt{k_n \log(k_n)} &\mbox{else.}
	\end{cases}
\end{equation}
and define further, for any $C > 0$
\begin{equation}\label{eq:def_K_C_set}
	\Kcal_C(k_n) = \left\{p=(x,y) \in \R \times \R_+ : \frac{k_n - C \kappa_n}{\xi_{\alpha,\nu}} \vee 0 \le e^{\frac{y}{2}}
	\le \frac{k_n + C \kappa_n}{\xi_{\alpha,\nu}} \wedge e^{R_n/2} \right\},
\end{equation}
\MS{What does $\vee$ and $\wedge$ mean?}

Note that if $k_n = \Omega(\log(n))$ then $e^{y/2} = \bigT{k_n}$ whenever $p=(x,y) \in \Kcal_C(k_n)$. The next lemma states that for a large class of functions $h(y)$, to compute the integral 
\[
	\int_{\Rcal_n} \rho(y,k_n) h(y) f_{\alpha,\nu}(x,y) \dd x \dd y
\]
it is enough to consider integration of $\Kcal_C(k_n)$ instead of $\Rcal_n$. More precisely, for any $\ell_n = (1 + \smallO{1})k_n$ it is enough to consider $\Kcal_C(\ell_n)$. The flexibility of using $(1+\smallO{1})k_n$ instead of $k_n$ will prove useful in later sections.

\begin{lemma}\label{lem:concentration_argument}
Let $\alpha > \frac{1}{2}$, $\nu > 0$, $\{k_n\}_{n \ge 1}$ be any positive sequence such that $k_n = o(n^{\frac{1}{2\alpha + 1}})$ and let $\ell_n = k_n(1 + \epsilon_n)$, with $\epsilon_n \to 0$. In addition let $\beta < \alpha$ and $h : \R_+ \rightarrow  \R$ be a any function such that $h(y) = g(y)e^{\beta y}$ with $|g(y)|$ uniformly bounded on $\R_+$. Then, if we define for $C > 0$
\[
	\lambda_n^\pm = \ell_n \pm C \kappa_n, 
	\quad a_n^\pm = 2 \log\left(\frac{\lambda_n^\pm}{\xi_{\alpha,\nu}}\right),
\] 
we have
\begin{equation}\label{eq:error_bound_int_rho_not_K}
\begin{aligned}
	\int_{\R_+ \setminus [a_n^-, a_n^+]} \rho(y,k_n) h(y) \alpha e^{-\alpha y} \dd x \dd y
	&= \begin{cases}
		\bigO{\log(n)^{k_n} n^{-\frac{C}{2}}} &\mbox{if } k_n = \bigT{1} \\
		\bigO{k_n^{-(1+C^2)/2}} &\mbox{else.}
	\end{cases}
\end{aligned}
\end{equation}
as $n \to \infty$.

@@

\subsection{Average degree in the hyperbolic random graph}\label{ssec:average_degree_HP_n}

Recall that under the coupling between the hyperbolic random graph and the finite box model $|x-x^\prime|_{\pi e^{R_n/2}} \le \Omega(r,r^\prime)$, while the coupling lemma (Lemma~\ref{lem:asymptotics_Omega_hyperbolic}) gives that  
\[
	e^{\frac{1}{2}(y+y^\prime)} - K e^{\frac{3}{2}(y+y^\prime) - R_n} \leq \Omega(r, r^\prime) 
		\leq  e^{\frac{1}{2}(y+y^\prime)} + K e^{\frac{3}{2}(y+y^\prime) - R_n},
\]
for $y + y^\prime < R_n$. This result enables us to determine the measure of a ball around a given point $p=(0,y)$ which will be fairly useful in our subsequent analysis.

@@

\begin{proof}
We will split the computation of $\Mu{\BallHyp{0,y}}$ into the cases $y^\prime > R_n - y$ and $y^\prime \le R_n - y$ where for the latter we utilize Lemma~\ref{lem:asymptotics_Omega_hyperbolic}. Recall that $\Mu{\BallPo{0,y}} = \xi_{\alpha,\nu} e^{y/2}$ where $\xi_{\alpha,\nu} = \frac{4\alpha\nu}{\pi(2\alpha - 1)}$.

@@

Next we will establish upper and lower bounds on $\Mu{\BallHyp{(0,y)} \cap\Rcal_n[(0,R_n-y)]}$. Using Lemma~\ref{lem:asymptotics_Omega_hyperbolic} we have 
\begin{align*}
	\Mu{\BallHyp{(0,y)} \cap\Rcal_n[(0,R_n-y)]} 
	&\le \frac{2\nu \alpha}{\pi} \int_{0}^{R_n - y} \left(e^{\frac{y + y^\prime}{2}} + K e^{\frac{3}{2}(y + y^\prime) - R_n}\right)
		e^{-\alpha y^\prime} \dd y^\prime\\
	&= \Mu{\BallPo{0,y}}\left(1 - e^{-(\alpha - \frac{1}{2})(R_n - y)}\right) \\
	&\hspace{10pt}+ \frac{2\nu \alpha}{\pi} K e^{\frac{3 y}{2} - R_n}\int_0^{R_n-y} e^{(\frac{3}{2} - \alpha)y^\prime} \dd y^\prime
\end{align*}
The last integral depends on the value of $\alpha$,
\[
	\int_0^{R_n-y} e^{(\frac{3}{2} - \alpha)y^\prime} \dd y^\prime
	= \begin{cases}
		\frac{2}{3 - 2\alpha}\left(e^{(\frac{3}{2} - \alpha)(R_n - y)} - 1\right) &\mbox{if } 1/2 < \alpha < 3/2,\\
		R_n - y &\mbox{if } \alpha = 3/2,\\
		\frac{2}{2\alpha-3}\left(1 - e^{-(\alpha - \frac{3}{2})(R_n - y)}\right) &\mbox{if } \alpha > 3/2.
	\end{cases}
\]
Therefore we get
\begin{align*}
	&\hspace{-30pt}\frac{2\nu \alpha}{\pi} K e^{\frac{3 y}{2} - R_n}
		\int_0^{R_n-y} e^{(\frac{3}{2} - \alpha)y^\prime} \dd y^\prime\\
	&= \Mu{\BallPo{0,y}}\begin{cases}
		\frac{(2\alpha - 1)K}{3 - 2\alpha}\left(e^{-(\alpha - \frac{1}{2})(R_n - y)} - e^{-(R_n - y)}\right)
		&\mbox{if } 1/2 < \alpha < 3/2,\\
		\frac{(2\alpha -1)K}{2} (R_n - y)e^{-(R_n - y)} &\mbox{if } \alpha = 3/2,\\
		\frac{(2\alpha - 1)K}{2\alpha - 3} \left(e^{-(R_n - y)} - e^{-(\alpha - \frac{1}{2})(R_n - y)}\right)
		&\mbox{if } \alpha > 3/2.
	\end{cases}\\
	&= \Mu{\BallPo{0,y}} \phi_{\H,n}^{(2)}(y)
\end{align*}
and hence
\[
	\Mu{\BallHyp{(0,y)} \cap\Rcal_n[(0,R_n-y)]} = \Mu{\BallPo{0,y}}\left(1 - e^{-(\alpha - \frac{1}{2})(R_n - y)} 
	+ \phi_{\H,n}^{(2)}(y)\right).
\]
Combining this with~\eqref{eq:mu_hyperbolic_ball_part_1} yields the required upper bound.

@@

\begin{lemma}\label{lem:average_degree_P_n}
For all $p \in \Rcal_n$, such that $y > 2\log(\pi/2)$,
\[
	\mu_{\alpha,\nu}(B_{\Pcal,n}(p)) = \mu_{\alpha,\nu}(B_\Pcal(p))\left(1 - \phi_n(y)\right)
\]
where $\phi_n(y) \ge 0$ is given by
\[
	\phi_n(y) = \left(\frac{\pi}{2}\right)^{-(2\alpha - 1)}e^{-(\alpha-\frac{1}{2})(R_n - y)}
	- \frac{(2\alpha - 1)\pi}{4\alpha}\left(\left(\frac{\pi}{2}\right)^{-2\alpha} 
	e^{-(\alpha - \frac{1}{2})(R_n - y)} - e^{-(\alpha - \frac{1}{2})R_n - \frac{y}{2}}\right).
\]
On the other hand, if $y \le 2 \log(\pi/2)$ then
\[
	\mu_{\alpha,\nu}(B_{\Pcal,n}(p)) = \mu_{\alpha,\nu}(B_\Pcal(p))\left(1 - e^{-(\alpha - \frac{1}{2})R_n}\right).
\]
\end{lemma}

@@

\subsection{Concentration argument for hyperbolic and finite box model}

With the results on the average degrees in the two finite Poisson models, we can now establish a result that allows us to extend the concentration argument to the case where instead of $\rho(y,k)$ we consider the degree distributions $\rho_{\HP,n}(y,k)$ and $\rho_{n}(y,k)$ in $G_{\HP,n}$ and $G_{\Pcal,n}$, respectively. The lemma is stated in a more general form, to make it applicable to slightly other cases later in the paper. To understand the general conditions presented in the statement of the lemma we write $\mu_n(y) = \Mu{\BallPon{y}}$ and $\mu(y) = \Mu{\BallPo{y}}$ and recall the result from Lemma~\ref{lem:average_degree_P_n}
\[
	\mu_n(y) = \mu(y)(1 - \phi_n(y)).
\]
For the error function $\phi_n(y)$ we note that for any $0 < \varepsilon < 1$, as $n \to \infty$, 
\[
	\sup_{0 \le y \le (1-\varepsilon)R_n} \phi_n(y) = \smallO{1},
\]
while
\[
	\frac{\partial \phi_n(y)}{\partial y} = \bigT{\phi_n(y)}.
\]
These are exactly the crucial properties of the error function that allow for a concentration argument. Moreover, if the integrand also contains a function $g_n(y)$ that is reasonably nice, we can replace the $\rho_{n}(y,k_n)$ in the integrand by $\rho(y,k_n)$. In particular, the triangle counting function $\Delta_{\Pcal}(y)$ will satisfy these conditions, see Section~\ref{sec:clustering_Pn_to_P}.

\begin{lemma}\label{lem:concentration_argument_rho_approximation}
Let $\alpha > \frac{1}{2}, \nu > 0$, $k_n$ be any positive sequence such that $k_n = o(n^{\frac{1}{2\alpha + 1}})$, $\ell_n = (1 + \epsilon_n)k_n$, with $\epsilon_n \to 0$ and let $\Kcal_{C}(\ell_n)$ be defined as in~\eqref{eq:def_K_C_set}. In addition, define $\hat{\rho}_n(y,k) = \Prob{\Po(\hat{\mu}_n(y)) = k}$, where $\hat{\mu}_n(y)$ is any function satisfying 
\[
	\Mu{\BallPo{0,y}}(1 - \phi_n^{(1)}(y)) \le \hat{\mu}_n(y) \le \Mu{\BallPo{0,y}}(1 + \phi_n^{(2)}(y))
\]
where $\phi_n^{(i)} : \R_+ \to \R_+$ are such that $\sup_{0 \le y \le (1-\varepsilon)R_n} \phi_n(y)^{(i)} = \smallO{1}$ for any $0 < \varepsilon < 1$ and $\partial \phi_n^{(i)}(y)/\partial y = \bigT{\phi_n^{(i)}(y)}$. Then the following holds for some $C > 0$ large enough:
\begin{enumerate}
\item for any sequence of functions $h_n(y) : \R_+ \mapsto \R$ such that $h_n(y) = \hat{\rho}_n(y,k_n) \bigO{k_n^s e^{\beta y}}$ for some $s < 2\alpha(2\alpha + 1)$ and $\beta < \alpha$, we have,
\[
	\int_{\Rcal_n} h_n(y) f_{\alpha,\nu}(x,y) \dd x \dd y 
	= (1 + o(1)) \int_{\Kcal_{C}(\ell_n)} h_n(y) f_{\alpha,\nu}(x,y) \dd x \dd y.
\]
\item If, in addition, $h_n(y) = \hat{\rho}_n(y,k_n) g_n(y)$ with $g_n$ satisfying $g_n(y(1+\smallO{1}) = (1+\smallO{1})g_n(y)$, we have,
\[
	\int_{\Rcal_n} h_n(y) f_{\alpha,\nu}(x,y) \dd x \dd y 
	= (1 + o(1)) \int_{\Rcal_n} \rho(y,k_n) g_n(y) f_{\alpha,\nu}(x,y) \dd x \dd y.
\]
\end{enumerate}
\end{lemma}

\begin{proof}
Similarly to the proof of Lemma~\ref{lem:concentration_argument} we define 
\[
	\lambda_n^\pm = \ell_n \pm C \kappa_n, \quad \text{and} \quad a_n^\pm = 2 \log\left(\frac{\lambda_n^\pm}{\xi_{\alpha,\nu}}\right).
\] 
In addition we fix $\varepsilon > 0$ such that $\varepsilon < \min\{1- \frac{s}{2\alpha(2\alpha + 1)}, 1\}$. Then since $h_n(y) = \hat{\rho}_n(y,k_n) \bigO{k_n^s}$ we have
\begin{align*}
	k_n^s \int_{(1-\varepsilon)R_n}^{R_n} \hat{\rho}_n(y,k_n) e^{-\alpha y} \dd y
	&= \bigO{1} \hat{\rho}_n((1-\varepsilon)R_n,k_n) k_n^s e^{-\alpha(1-\varepsilon)R_n} \\
	&= \bigO{\hat{\rho}_n((1-\varepsilon)R_n,k_n) k_n^s n^{-2\alpha(1 - \varepsilon)}} \\
	&= \smallO{ \hat{\rho}_n((1-\varepsilon)R_n,k_n) n^{\frac{s}{2\alpha + 1} - 2\alpha(1-\varepsilon)}} = \smallO{1},
\end{align*}
where the last step follows by our choice of $\varepsilon$. Hence it is enough to prove both statement on $\Rcal_n((1-\varepsilon)R_n, R_n)$ instead of $\Rcal_n$.
\PvdH{This is where we use that $s < 2\alpha(2\alpha + 1)$. I think this can be removed by using a more careful bound but I am not sure this is needed.}

@@

In particular we conclude that, similarly to the infinite limit model, concentration arguments as described in Remark~\ref{rmk:concentration_argument} can be applied in the case of the hyperbolic random graph and finite box model.

### 2018-11-16 introduction.tex
typos in the proof of thm 1.1. in section 2 and typos in section 3

Our first result shows that the local clustering coefficient of the hyperbolic random graph converges in expectation to a constant $c_\infty$, which we can explicitly write down in terms of the special functions defined above.

>>>>

Our first result shows that the local clustering coefficient of the hyperbolic random graph converges in expectation to a constant $c_\infty$, which we can explicitly write down in terms of the special functions mentioned above.
